---
title: "Bronez Approach"
author: "Amanda Koepke"
date: "2023-06-28"
output: 
  pdf_document: 
    fig_crop: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RSpectra) #eigensolving
library(geigen)
library(tidyverse)
library(fields)
```

The data consists of observations $x_i$ and time points $t_i$ for $i \in 1, \ldots,N$. The time points are not necessarily evenly spaced, but we will treat them as evenly spaced in the example below. 

We define $K$, the number of tapers we want to use, and $\omega$ (f.w in code), the bandwidth. In general a good choice of bandwidth is $4/N$. 

To calculate the tapers, we don't need the $x_i$, just the symmetric distance matrix $\boldsymbol{\Delta}$, where $\Delta_{n,m}=|t_n-t_m|$. 


## Bronez approach

Extending multitaper spectral analysis to irregularly spaced data, Bronez proposed a spectral estimate that is the 
<!-- average spectral power  -->
integrated spectrum over a specified bandwidth. 

It's a quadratic spectral estimator
\[
\bar S_\omega=\mathbf{x}^*\mathbf{Q}_\omega\mathbf{x},
\]
where $\mathbf{Q}_\omega=\mathbf{w}_\omega \mathbf{w}_\omega^*$. To solve for these $\mathbf{w}$ for a given $\omega$ and process half-bandwidth $\beta$, we solve a generalized eigenvalue problem described below. 

<!-- Under the Bronez approach,  -->
First we create matrices $\mathbf{A}$ and $\mathbf{B}$. 

The matrix $\mathbf{A}$ is defined as 

\begin{eqnarray*}
A_{nm} = \int_{f_c-\omega}^{f_c+\omega}e^{i 2 \pi f (t_n-t_m) } df &=& \int_{f_c-\omega}^{f_c+\omega}e^{i 2 \pi f \Delta_{n,m} } df \\
&=&  \frac{i (e^{i 2 \pi \Delta_{n,m} (f_c-\omega)}-e^{i 2 \pi \Delta_{n,m} (f_c+\omega)})}{2 \pi \Delta_{n,m}} \\
&=&  \frac{e^{i 2 \pi \Delta_{n,m} f_c}}{\pi \Delta_{n,m}} \frac{e^{i 2 \pi \Delta_{n,m} \omega}-e^{-i 2 \pi \Delta_{n,m} \omega}}{2i} \\
&=&  \frac{e^{i 2 \pi \Delta_{n,m} f_c}}{\pi \Delta_{n,m}} sin(2 \pi \Delta_{n,m} \omega)
\end{eqnarray*}

When $n=m$, $A_{nm} = \int_{f_c-\omega}^{f_c+\omega}e^{0} df = 2\omega$.

This matrix depends on the center frequency, $f_c$. 

The $\mathbf{B}$ matrix is defined as 
\[
B_{nm}=\int_{-\beta}^{\beta}e^{i 2 \pi f (t_n-t_m) } df
\]

When $t_n=t_m$, $B_{nn}=\int_{-\beta}^{\beta}e^{0} df = 2\beta$. When $\beta=1/2$, as we assume, $B_{nn}=1$.

When $t_n \neq t_m$, $B_{nm}=\frac {\mbox{sin} \left(2 \beta \pi (t_n-t_m) \right)}{\pi (t_n-t_m)}$, which reduces to $\frac {\mbox{sin} \left(\pi (t_n-t_m) \right)}{\pi (t_n-t_m)}$ when $\beta=1/2$. When the times are all integer values, as they are in this application, $B_{nm}=0$, so $\mathbf{B}$ simplifies further to the identity matrix. 

Then to determine the tapers for a given frequency, we solve the eigenvalue problem
\[
\mathbf{A} \underline w_k = \lambda_k \mathbf{B} \underline w_k
\]

where $\underline w_k$ is the $k^{th}$ eigenvector. When choosing the number of tapers $K$, ensure the eigenvalues are close to 1. 

Under the Bronez approach, $\mathbf{A}$ varies by center frequency ($f_c$, f.c in the code) and this eigenvalue problem must be solved for each frequency. Typically we solve this for $\lfloor N/2 \rfloor+1$ frequencies. This can be computationally expensive, and because of this Bronez's method did not seem to catch on. In our code, we parallelize the computation to speed it up, though this approach is still quite computationally expensive for the size of datasets we expect to see. 

### Calculate spectral estimate

With the tapers, we can estimate the spectrum using
\[
\frac{1}{K}\sum_{k=1}^{K} \left|\underline w_k^* \underline x \right|^2 =\frac{1}{K}\sum_{k=1}^{K} \left| \sum_{i=1}^N w_{k,i} x_i \right|^2
\]

### Variance

Bronez showed that $Var(\hat{P}(\mathcal A))=||w^*(\mathcal A)R_xw(\mathcal A)||^2$, where $\mathcal A$ defines the interval over which the spectral power is being averaged, $f_c \pm \omega$.

We have generalized this to show $Cov(\hat{P}(\mathcal A_i),\hat{P}(\mathcal A_j))=||\underline w^*(\mathcal A_i) R_x \underline w(\mathcal A_j)||^2$, where $R_x$ is the $NxN$ positive semidefinite symmetric autocovariance matrix whose $(n,m)^{th}$ element is $r_x(\Delta_{n,m})$.

For white noise, $R_x=I$, so for each pair of center frequencies $(i,j)$ we'd calculate $||\underline w^*(\mathcal A_i) \underline w(\mathcal A_j)||^2$


## Chave approach

Using Bronez's approach, tapers must be calculated for each frequency, which is computationally expensive, in addition to numerical issues encountered solving the eigenvalue problem at the extremes of the process band. Chave proposed an alternative spectral estimator, which redefines 
\[
\underline w_k=\underline v_k \mbox{exp}[-i2 \pi f_c \Delta_{n,m}],
\]
giving the new eigenvalue problem 
\[
\mathbf{A}' \underline v_k=\lambda_k \mathbf{B}' \underline v_k.
\]

\textit{(we are still missing some steps in the derivation here)}

Here
\[
A'_{nm} = \frac{\mbox{sin}(2 \pi \Delta_{n,m} \omega)}{\pi \Delta_{n,m}}, \]
\[
A_{nn} = 2\omega,
\]
and 
\[
B'_{nm} = e^{-i 2 \pi f_c \Delta_{n,m}} \frac{\mbox{sin} 2 \pi \beta \Delta_{n,m}}{\pi \Delta_{n,m}},
\]
which reduces to the identity matrix when $\beta=1/2$ and $\Delta_{n,m}=t_n-t_m$ is an integer. 

This leads to the spectral estimator

\[
\frac{1}{K}\sum_{k=1}^{K} \left| \sum_{i=1}^N v_{k,i} x_i e^{-i2\pi f_c t_i} \right|^2.
\]

### Variance

Using a similar form to the Bronez approach, we can estimate a covariance.
\[
Cov(i,j)=||\underline w^*(f_i) R_x \underline w(f_j)||^2
\]
where $\underline w=\underline v \mbox{exp}[-i2 \pi f_c \underline{t}]$ and $\underline{t}$ is the vector of timepoints. 

For white noise, $R_x=I$, so for each frequency pair $(i,j)$ we'd calculate 

\[
Cov(i,j)=|| (\underline v e^{-i2 \pi f_i \underline{t}})^* (\underline v e^{-i2 \pi f_j \underline{t}} )||^2
\]

<!-- \[ -->
<!-- Cov(\hat{P}(A_1),\hat{P}(A_2))=||w^*(A_1) w(A_2)||^2=|| (v e^{-i2 \pi f \mathbf{t}})^* (v e^{-i2 \pi f \mathbf{t}} )||^2 -->
<!-- \] -->

<!-- for(i in 1:N.fourier){ -->
<!-- j = 1 -->
<!--   while(j <= i){ -->
<!--     Cov.mat_chave[i,j] <- norm(Conj(t(test_chave$tapers*exp(-im*2*pi*freq[i]*1:N)))%*%(test_chave$tapers*exp(-im*2*pi*freq[j]*1:N)), type = "2") -->
<!--     j = j+1 -->
<!--   } -->
<!-- } -->


## Bronez approach code

```{r,eval=FALSE,tidy='styler', tidy.opts=list(strict=FALSE)}

################################################################################################
## tapers used later are calculated by frequency in BronezTapers.R, code here for completeness 
################################################################################################

### calculate eigenvalues and eigenvectors, return the K largest in magnitude (constrained to 
### be greater than zero, so will be last K)
get_k_geigen=function(Ra,Rb,thek,f.w){
  evs=geigen(Ra,Rb, symmetric = TRUE)

  N=dim(evs$vectors)[1]
  return(list("weights" = evs$vectors[,(N-thek + 1):N], 
              "eigenvalues" = sort(evs$values, decreasing = TRUE)[1:thek]))
}

### calculate tapers using Bronez approach
get.weights_bronez <- function(input.list){
  t.n = input.list$t.n
  K = input.list$K
  f.c = input.list$f.c
  f.w = input.list$f.w
  dist.mat = input.list$tn_m
  index = input.list$index
  print(index)

  R.b <- diag(1,nrow = length(t.n),ncol = length(t.n))

  j = complex(real = 0, imaginary = 1)
  R.a <- exp(j*2*pi*f.c*dist.mat)*(sin(2*pi*f.w*(dist.mat))/(pi*dist.mat))
  R.a[row(R.a) == col(R.a)] <- f.w*2

  #Solve the eigenvalue problem
  out <- get_k_geigen(R.a,R.b,K,f.w)

  return(out)
}

### create equally spaced time intervals from 1 to N
N <- 256
N.fourier <- floor(N/2) + 1
freq <- seq(0,0.5, length.out = N.fourier)

input.list=list()
for(i in 1:length(freq)){
  input.list[[i]]=list("t.n"=1:N,
                       "K"=3,
                       "f.c"=freq[i],
                       "f.w"=4/N,
                       "tn_m"=rdist(1:N),
                       "index" = i)
}

##################################################################################################
# The below parallelizes the code and creates a list where each element corresponds to a frequency
##################################################################################################
startTime = Sys.time()
parResult = mclapply(input.list, get.weights_bronez, mc.cores = 3)
print(Sys.time()-startTime)

saveRDS(parResult,"/home/aak3/NIST/ClockDataAnalysis/Code/BronezResults256_070423.Rds")

```

### Read in the tapers generated by the above code

```{r,fig.width=5,fig.height=3}

#read in what was written above
tapers256=readRDS("/home/aak3/NIST/ClockDataAnalysis/Code/BronezResults256_070423.Rds")

taper.mat <- tapers256
K=dim(taper.mat[[1]]$weights)[2]
N = length(taper.mat[[1]]$weights[,1])
N.fourier = floor(N/2) + 1
freq = seq(0,0.5, length.out = N.fourier)
set.seed(1990)
x.t <- rnorm(N,0,1)

### check eigenvalues
eigsDF=data.frame(freq=1,k=1:K,eigs=taper.mat[[1]]$eigenvalues)
for(i in 1:length(taper.mat)){
  eigsDF=bind_rows(eigsDF,data.frame(freq=i,k=1:K,eigs=taper.mat[[i]]$eigenvalues))
  # (taper.mat[[i]]$eigenvalues)
}

ggplot(eigsDF,aes(k,eigs))+
  geom_point()

##eigenvalues are close to 1



## look at a tapers
centerFreqNum=1

plotTapers=data.frame(freq=centerFreqNum,k=1,n=1:N, weight=Re(tapers256[[centerFreqNum]]$weights[,1]))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=2,n=1:N,weight=Re(tapers256[[centerFreqNum]]$weights[,2])))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=3,n=1:N,weight=Re(tapers256[[centerFreqNum]]$weights[,3])))
plotTapers$k =as.factor(plotTapers$k)

ggplot(plotTapers,aes(n,weight,col=k))+
  geom_line()


centerFreqNum=15

plotTapers=data.frame(freq=centerFreqNum,k=1,n=1:N, weight=Re(tapers256[[centerFreqNum]]$weights[,1]))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=2,n=1:N,weight=Re(tapers256[[centerFreqNum]]$weights[,2])))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=3,n=1:N,weight=Re(tapers256[[centerFreqNum]]$weights[,3])))
plotTapers$k =as.factor(plotTapers$k)

ggplot(plotTapers,aes(n,weight,col=k))+
  geom_line()

```

### Calculate spectral estimate

```{r}
spec.hat <- rep(NA, times = N.fourier)

for(i in 1:length(freq)){
  # print(i)
  spec.hat[i] <- (1/K)*sum(abs(Conj(t(taper.mat[[i]]$weights))%*%x.t)^2)
}

```


<!-- ```{r} -->
<!-- plot(freq, spec.hat)  -->
<!-- abline(h=1) #variance of data -->

<!-- ``` -->


## Variance


```{r}
Cov.mat_bronez <- matrix(NA, nrow = N.fourier, ncol = N.fourier)

for(i in 1:N.fourier){
  j = 1
  while(j <= i){
    Cov.mat_bronez[i,j] <- norm(Conj(t(tapers256[[i]]$weights))%*%(tapers256[[j]]$weights), type = "2") 
    j = j+1
  }
}

Cov.mat_bronez[upper.tri(Cov.mat_bronez)] <- t(Cov.mat_bronez)[upper.tri(Cov.mat_bronez)]

# plot(freq, spec.hat) 
specDF=data.frame(freq=freq,spec.hat=spec.hat,variance=diag(Cov.mat_bronez))
ggplot(specDF,aes(freq,spec.hat,ymin=spec.hat-2*sqrt(variance),ymax=spec.hat+2*sqrt(variance)))+
  geom_point()+
  geom_errorbar()+
  geom_hline(yintercept = 1)
```

## Chave approach


```{r,message=FALSE}
##compare to Chave estimate
source("/home/aak3/NIST/ClockDataAnalysis/Code/SA_ImportantFunctions.R")
test_chave <- multitaper_est(X.t = x.t, W = 4/N, K = 3)

# plot(freq, test_chave$spectrum, col = "blue")
# points(freq, spec.hat) 
# plot(freq, abs(test_chave$spectrum - spec.hat))

Cov.mat_chave <- matrix(NA, nrow = N.fourier, ncol = N.fourier)

for(i in 1:N.fourier){
  j = 1
  while(j <= i){
    Cov.mat_chave[i,j] <- norm(Conj(t(test_chave$tapers*exp(-im*2*pi*freq[i]*1:N)))%*%(test_chave$tapers*exp(-im*2*pi*freq[j]*1:N)), type = "2") 
    j = j+1
  }
}

Cov.mat_chave[upper.tri(Cov.mat_chave)] <- t(Cov.mat_chave)[upper.tri(Cov.mat_chave)]

Cov.mat_bronez[1:5,1:5]
Cov.mat_chave[1:5,1:5]
norm(Cov.mat_bronez - Cov.mat_chave, type = "2")

specDF_bronez=data.frame(freq=freq,spec.hat=spec.hat,variance=diag(Cov.mat_bronez),type="Bronez")
specDF_chave=data.frame(freq=freq,spec.hat=test_chave$spectrum,variance=diag(Cov.mat_chave),type="Chave")
specDF_both=bind_rows(specDF_bronez,specDF_chave)

ggplot(specDF_both,aes(freq,spec.hat,ymin=spec.hat-2*sqrt(variance),ymax=spec.hat+2*sqrt(variance),col=type))+
  geom_point()+
  geom_errorbar()+
  geom_hline(yintercept = 1)


```









# for data with gaps

```{r,eval=FALSE}
library(RSpectra) #eigensolving
library(geigen)
library(tidyverse)
library(fields)
library(parallel)
################################################################################################
## tapers used later are calculated by frequency in BronezTapers.R, code here for completeness
################################################################################################

### calculate eigenvalues and eigenvectors, return the K largest in magnitude (constrained to
### be greater than zero, so will be last K)
get_k_geigen=function(Ra,Rb,thek,f.w){
  evs=geigen(Ra,Rb, symmetric = TRUE)

  N=dim(evs$vectors)[1]
  return(list("weights" = evs$vectors[,(N-thek + 1):N],
              "eigenvalues" = sort(evs$values, decreasing = TRUE)[1:thek]))
}

### calculate tapers using Bronez approach
get.weights_bronez <- function(input.list){
  t.n=input.list$t.n
  K=input.list$K
  f.c=input.list$f.c
  f.w=input.list$f.w
  dist.mat = input.list$tn_m
  index = input.list$index
  print(index)

  R.b <- diag(1,nrow = length(t.n),ncol = length(t.n))

  j = complex(real = 0, imaginary = 1)
  R.a <- exp(j*2*pi*f.c*dist.mat)*(sin(2*pi*f.w*(dist.mat))/(pi*dist.mat))
  R.a[row(R.a) == col(R.a)] <- f.w*2

  #Solve the eigenvalue problem
  out <- get_k_geigen(R.a,R.b,K,f.w)

  return(out)
}


###gaps####
t.vec <- 1:300 #time vector
x.t <- rnorm(300) #data
t.vec[c(20:35,50:63, 100:130)] <- NA #take out values
x.t[c(20:35,50:63, 100:130)] <- NA #take out values
t.vec <- na.omit(t.vec) #vector of times with data
dist.mat <- rdist(t.vec) #distance matrix (delta_nm)

N <- length(t.vec)
N.fourier <- floor(N/2) + 1
freq <- seq(0,0.5, length.out = N.fourier)


input.list=list()
for(i in 1:length(freq)){
  input.list[[i]]=list("t.n"=t.vec,
                       "K"=3,
                       "f.c"=freq[i],
                       "f.w"=4/N,
                       "tn_m"=dist.mat,
                       "index" = i)
}

##################################################################################################
# The below parallelizes the code and creates a list where each element corresponds to a frequency
##################################################################################################
startTime = Sys.time()
parResult = mclapply(input.list, get.weights_bronez, mc.cores = 3)
print(Sys.time()-startTime)

saveRDS(parResult,"/home/aak3/NIST/ClockDataAnalysis/Code/BronezResults256_gaps_070523.Rds")

```


```{r}
rm(list=ls())

# this is how I simulated data in the above chunk, running here since the above isn't executed in this file

###gaps####
t.vec <- 1:300 #time vector
x.t <- rnorm(300) #data
omitted<-c(20:35,50:63, 100:130)
t.vec[c(20:35,50:63, 100:130)] <- NA #take out values
x.t[c(20:35,50:63, 100:130)] <- NA #take out values
t.vec <- na.omit(t.vec) #vector of times with data
dist.mat <- rdist(t.vec) #distance matrix (delta_nm)

N <- length(t.vec)
N.fourier <- floor(N/2) + 1
freq <- seq(0,0.5, length.out = N.fourier)

```

### Read in the tapers generated by the above code

```{r,fig.width=5,fig.height=3}

#read in what was written above
tapers256=readRDS("/home/aak3/NIST/ClockDataAnalysis/Code/BronezResults256_gaps_070523.Rds")

taper.mat <- tapers256
K=dim(taper.mat[[1]]$weights)[2]
# N = length(taper.mat[[1]]$weights[,1])
# N.fourier = floor(N/2) + 1
# freq = seq(0,0.5, length.out = N.fourier)
# set.seed(1990)
# x.t <- rnorm(N,0,1)


### check eigenvalues

eigsDF=data.frame(freq=1,k=1:K,eigs=taper.mat[[1]]$eigenvalues)
for(i in 1:length(taper.mat)){
  eigsDF=bind_rows(eigsDF,data.frame(freq=i,k=1:K,eigs=taper.mat[[i]]$eigenvalues))
  # (taper.mat[[i]]$eigenvalues)
}

ggplot(eigsDF,aes(k,eigs))+
  geom_point()

#eigenvalues are close to 1

## look at a tapers
centerFreqNum=1

plotTapers=data.frame(freq=centerFreqNum,k=1,n=t.vec, weight=Re(taper.mat[[centerFreqNum]]$weights[,1]))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=2,n=t.vec,weight=Re(taper.mat[[centerFreqNum]]$weights[,2])))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=3,n=t.vec,weight=Re(taper.mat[[centerFreqNum]]$weights[,3])))

plotTapersOmitted1=data.frame(freq=centerFreqNum,k=1,n=omitted,weight=NA)
plotTapersOmitted2=data.frame(freq=centerFreqNum,k=2,n=omitted,weight=NA)
plotTapersOmitted3=data.frame(freq=centerFreqNum,k=3,n=omitted,weight=NA)

plotTapers=bind_rows(plotTapers,plotTapersOmitted1,plotTapersOmitted2,plotTapersOmitted3)
                     
plotTapers$k =as.factor(plotTapers$k)

ggplot(plotTapers,aes(n,weight,col=k))+
  geom_line()


centerFreqNum=15

plotTapers=data.frame(freq=centerFreqNum,k=1,n=t.vec, weight=Re(taper.mat[[centerFreqNum]]$weights[,1]))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=2,n=t.vec,weight=Re(taper.mat[[centerFreqNum]]$weights[,2])))
plotTapers=bind_rows(plotTapers,data.frame(freq=centerFreqNum,k=3,n=t.vec,weight=Re(taper.mat[[centerFreqNum]]$weights[,3])))

plotTapersOmitted1=data.frame(freq=centerFreqNum,k=1,n=omitted,weight=NA)
plotTapersOmitted2=data.frame(freq=centerFreqNum,k=2,n=omitted,weight=NA)
plotTapersOmitted3=data.frame(freq=centerFreqNum,k=3,n=omitted,weight=NA)

plotTapers=bind_rows(plotTapers,plotTapersOmitted1,plotTapersOmitted2,plotTapersOmitted3)
                     
plotTapers$k =as.factor(plotTapers$k)

ggplot(plotTapers,aes(n,weight,col=k))+
  geom_line()

```


### Calculate spectral estimate

With the tapers, we can estimate the spectrum using
\[
\frac{1}{K}\sum_{k=1}^{K} \left|\underline w_k^* \underline x \right|^2 =\frac{1}{K}\sum_{k=1}^{K} \left| \sum_{i=1}^N w_{k,i} x_i \right|^2
\]

```{r}
spec.hat <- rep(NA, times = N.fourier)

for(i in 1:length(freq)){
  # print(i)
  spec.hat[i] <- (1/K)*sum(abs(Conj(t(taper.mat[[i]]$weights))%*%as.matrix(na.omit(x.t),ncol=1))^2)
}

```

<!-- ## Variance -->

<!-- Bronez showed that $Var(\hat{P}(A))=||w^*(A)R_xw(A)||^2$. -->

<!-- Mike has generalized this to show $Cov(\hat{P}(A_1),\hat{P}(A_2))=||w^*(A_1) R_x w(A_2)||^2$, where $R_x$ is the autocovariance matrix. -->


<!-- For white noise, $R_x=I$, so for each frequency we'd calculate $||w^*(A_1) w(A_2)||^2$ -->


```{r}
Cov.mat_bronez <- matrix(NA, nrow = N.fourier, ncol = N.fourier)

for(i in 1:N.fourier){
  j = 1
  while(j <= i){
    Cov.mat_bronez[i,j] <- norm(Conj(t(taper.mat[[i]]$weights))%*%(taper.mat[[j]]$weights), type = "2")
    j = j+1
  }
}

Cov.mat_bronez[upper.tri(Cov.mat_bronez)] <- t(Cov.mat_bronez)[upper.tri(Cov.mat_bronez)]

specDF=data.frame(freq=freq,spec.hat=spec.hat,variance=diag(Cov.mat_bronez))
ggplot(specDF,aes(freq,spec.hat,ymin=spec.hat-2*sqrt(variance),ymax=spec.hat+2*sqrt(variance)))+
  geom_point()+
  geom_errorbar()+
  geom_hline(yintercept = 1)

```



```{r,message=FALSE}
##compare to Chave estimate
source("/home/aak3/NIST/ClockDataAnalysis/Code/SA_ImportantFunctions.R")
test_chave <- multitaper_est(X.t = x.t, W = 4/N, K = 3)

# plot(freq, test_chave$spectrum, col = "blue")
# points(freq, spec.hat)
# plot(freq, abs(test_chave$spectrum - spec.hat))

Cov.mat_chave <- matrix(NA, nrow = N.fourier, ncol = N.fourier)

for(i in 1:N.fourier){
  j = 1
  while(j <= i){
    Cov.mat_chave[i,j] <- norm(Conj(t(test_chave$tapers*exp(-im*2*pi*freq[i]*t.vec)))%*%(test_chave$tapers*exp(-im*2*pi*freq[j]*t.vec)), type = "2")
    j = j+1
  }
}

Cov.mat_chave[upper.tri(Cov.mat_chave)] <- t(Cov.mat_chave)[upper.tri(Cov.mat_chave)]

Cov.mat_bronez[1:5,1:5]
Cov.mat_chave[1:5,1:5]
norm(Cov.mat_bronez - Cov.mat_chave, type = "2")


specDF_bronez=data.frame(freq=freq,spec.hat=spec.hat,variance=diag(Cov.mat_bronez),type="Bronez")
specDF_chave=data.frame(freq=freq,spec.hat=test_chave$spectrum,variance=diag(Cov.mat_chave),type="Chave")
specDF_both=bind_rows(specDF_bronez,specDF_chave)

ggplot(specDF_both,aes(freq,spec.hat,ymin=spec.hat-2*sqrt(variance),ymax=spec.hat+2*sqrt(variance),col=type))+
  geom_point()+
  geom_errorbar()+
  geom_hline(yintercept = 1)
```