---
title: "Multitaper Spectral Analysis on Data with Dead Time"
output: 
  html_document: 
    number_sections: true
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(arfima)
library(tidyverse)
library(fields)
library(multitaper)
library(geigen)
library(RSpectra)
library(R.utils)

load("C:/Users/cmb15/OneDrive - UCB-O365/NIST/ClockDataAnalysis/Data/ClockData.RData")

#functions:
h.k.prime <- function(k,N,N.prime){
  t <- 0:(N-1)
  h.k.N <- (2/(N+1))^(1/2)*sin((k + 1)*pi*(t + 1)/(N + 1)) #value of h' taper for t = 0, .., N-1
  if(N.prime > N){
    return(c(h.k.N, rep(0, times = N.prime-N)))#add in the 0's to pad the end to get up to 4096
  }
  else{
    return(h.k.N)
  }
}

avar_fn=function(y,tau){
  n=length(y)
  
  div=seq(1,n,by = tau)
  
  M=length(div)-1 #number of groups
  
  groupmeans = numeric(M)
  for(i in 1:M){
    groupmeans[i]=mean(y[div[i]:(div[i+1]-1)])
  }
  
  1/(2*(M-1)) * sum(diff(groupmeans)^2)
}

getAvars=function(N, y,taus){
  
  avars=numeric(length(taus))
  
  for (i in 1:length(taus)){
    avars[i]=avar_fn(y,taus[i])
  }
  
  m1=data.frame(taus=taus,avars=avars)  
  fit=lm(log(sqrt(avars))~log(taus),data = m1)
  slope=as.numeric(fit$coefficients[2])
  int=as.numeric(fit$coefficients[1])
  
  avarRes=data.frame(taus=taus,avars=avars,N=N,slope=slope,int=int)  
  
  ##########################################
  # get SE
  ##########################################
  SEests=data.frame()
  
  new=data.frame(N=N,out=exp(int+slope*log(N)), type="AD")
  new2=data.frame(N=N,out=sd(y)/sqrt(N), type="SE")
  new3=data.frame(N=N,out=1/sqrt(N), type="true")
  SEests=bind_rows(new,new2,new3)
  
  return(list(avarRes=avarRes,SEests=SEests))
}

multitaper_est <- function(X.t, NW, K){
  X.t <- X.t - mean(X.t, na.rm = TRUE) #demean
  N.long <- length(X.t)
  t.n <- 1:N.long
  missing.indices <- which(is.na(X.t))
  t.n[which(is.na(X.t))] <- NA


  W <- NW/length(na.omit(t.n))
  dist.mat <- rdist(na.omit(t.n))
  
  #create the A' matrix (Chave 2019 equation (22))
  A.prime <- (1/(pi*dist.mat))*sin(2*pi*W*dist.mat)
  A.prime[row(A.prime) == col(A.prime)] <- W*2
  eigdec <- eigs_sym(A.prime, k = K, which = "LM")
  
  
  eig_vecs <- eigdec$vectors #get only the vectors
  
  if(K ==1){
    if (mean(Re(eig_vecs))<0){
      eig_vecs <- -eig_vecs
    }
  }
  
  if(K >1){
    if(K >=2){
    if (mean(Re(eig_vecs[,1]))<0){
      eig_vecs[,1] <- -eig_vecs[,1]
    }
    if (Re(eig_vecs[2,2] - eig_vecs[1,2])<0){
      eig_vecs[,2] <- -eig_vecs[,2]
    }
    }
    if(K == 3){
      if (mean(Re(eig_vecs[,3]))<0){
      eig_vecs[,3] <- -eig_vecs[,3]
    }
    }
  }
  if(K >=4){
  #some sign maintenance
  for(i in seq(1,K,by = 2)){
    if (mean(Re(eig_vecs[,i]))<0){
      eig_vecs[,i] <- -eig_vecs[,i]
    }
  }
  
  for(i in seq(2,K-1,by = 2)){
    if (Re(eig_vecs[2,i] - eig_vecs[1,i])<0){
      eig_vecs[,i] <- -eig_vecs[,i]
    }
  }
  }
  
  ##use tapers to generate spectral estimate
  N <- length(na.omit(t.n))
  S.x.hat_MD <- rep(NA, times = floor(N/2) + 1)
  
  for(j in 0:floor(N/2)){
    k.vec <- rep(NA,times = K)
    for(k in 0:(K-1)){
      W.t <- eig_vecs[,k+1]*na.exclude(X.t)
      inner.sum <- sum(W.t*exp(-complex(real = 0, imaginary = 1)*2*pi*na.omit(t.n)*j/N), na.rm = TRUE)
      k.vec[k + 1] <- abs(inner.sum)^2
    }
    S.x.hat_MD[j+1] <- mean(k.vec)
  }
  
  
  return(list("tapers" = eig_vecs, "e.values" = eigdec$values, "spectrum" = S.x.hat_MD))
}


```

## The Data

First we show an example of data generated from an ARFIMA(0,d,0) process with d = 0.4.

This process follows:

$$\nabla^d X_t = \epsilon_t$$


where we let $d = 0.4$ and $var\{\epsilon_t\} = 1$. We show the full process and the dead time process where there are chunks of the data removed.

```{r, echo = FALSE}
N <- 2048
d <- 0.4
X.t <- X.t_missing <- arfima.sim(N,model = list(dfrac = d))

endpoints <- c(450,700,1000,1200)#end points of missing chunks

X.t_missing[c(endpoints[1]:endpoints[2], endpoints[3]:endpoints[4])] <- NA


plot(X.t,type = "l", main = "Full Data", ylab = "X(t)", xlab = "t")
plot(X.t_missing,type = "l", main = "Partial Data", ylab = "X(t)", xlab = "t")
```




## Mutltitaper Spectral Estimate (MTSE)

We find the MTSE of the full and partial data using K = 5 tapers.The spectrum of an ARFIMA process is a well known relationship, we show the log-log relationship here: $\log S(f) = \log(\sigma_\epsilon^2) - 2d\log(2|\sin\pi f|)$.

For the data above, recall that $d = 0.4$ and $\sigma_\epsilon^2 = 1$, so we plot this relationship along with the spectral estimates of the above data.

```{r,echo = FALSE}
MTSE_fd <- multitaper_est(X.t = X.t, NW = 2, K = 5)
MTSE_fd_missing <- multitaper_est(X.t = X.t_missing, NW = 2, K = 5)


f <- log10(seq(0,0.5,length.out = length(MTSE_fd$spectrum)))[-1]
f.missing <- log10(seq(0,0.5,length.out = length(MTSE_fd_missing$spectrum)))[-1]

plot(f,log10(MTSE_fd$spectrum[-1]), type = "l", main = "Full vs. Partial Data MTSE comparison, d = 0.4", xlab = "log10(f)")
lines(f.missing, log10(MTSE_fd_missing$spectrum[-1]), col = "red")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue")
legend(x = -3, y = 0.5, legend = c("Full Data", "Missing Data", "Truth"), col = c("black", "red", "blue"), lty = 1)


```


## Another Example

```{r}
N <- 2048
d <- 0.25
X.t <- X.t_missing <- arfima.sim(N,model = list(dfrac = d))

endpoints <- c(450,700,1000,1200)#end points of missing chunks

X.t_missing[c(endpoints[1]:endpoints[2], endpoints[3]:endpoints[4])] <- NA

MTSE_fd <- multitaper_est(X.t = X.t, NW = 2, K = 5)
MTSE_fd_missing <- multitaper_est(X.t = X.t_missing, NW = 2, K = 5)


f <- log10(seq(0,0.5,length.out = length(MTSE_fd$spectrum)))[-1]
f.missing <- log10(seq(0,0.5,length.out = length(MTSE_fd_missing$spectrum)))[-1]

plot(f,log10(MTSE_fd$spectrum[-1]), type = "l", main = "Full vs. Partial Data MTSE comparison, d = 0.25", ylab = "log10(S(f))", xlab = "log10(f)")
lines(f.missing, log10(MTSE_fd_missing$spectrum[-1]), col = "red")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue")
legend(x = -3, y = 0.5, legend = c("Full Data", "Missing Data", "Truth"), col = c("black", "red", "blue"), lty = 1)
```


## Finding $K$ and $W$

The above examples choose a $K$ and $W$ somewhat arbitrarily, although with some basis in previous work. According to (Percival, "Characterization of frequency stability..." 1991), for time and frequency data, typical values of $W$ are given by $\eta/N$ where $1 \leq \eta < 4$. The spectral windows of these estimators have a central lobe with approximate width $\eta/N$ for $\tau_1 = 1$ where $\tau_1$ is the time resolution of the data.

```{r}
MTSE_fd$e.values
MTSE_fd_missing$e.values
```
These seem to drop off faster than I would expect..I have to look into this.

According to literature the first $2NW$ eigenvalues are supposed to be close to 1. 

### Choosing W:

```{r}
MTSE_fd1 <- multitaper_est(X.t = X.t, NW = 1, K = 2)
MTSE_fd2 <- multitaper_est(X.t = X.t, NW = 2, K = 4)
MTSE_fd3 <- multitaper_est(X.t = X.t, NW = 3, K = 6)
MTSE_fd4 <- multitaper_est(X.t = X.t, NW = 4, K = 8)

MTSE_fd1$e.values
MTSE_fd2$e.values
MTSE_fd3$e.values
MTSE_fd4$e.values

```
it seems like 2NW - 1 are close to 1.

This might be a good choice for $K$, then $K = 2NW - 1$

```{r}

par(mfrow = c(2,2))

f <- log10(seq(0,0.5,length.out = length(MTSE_fd1$spectrum)))[-1]

plot(f,log10(MTSE_fd1$spectrum[-1]), type = "l", main = "NW = 1", ylab = "log10(S(f))", xlab = "log10(f)")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue", xlab = "log10(f)")
plot(f,log10(MTSE_fd2$spectrum[-1]), type = "l", main = "NW = 2", ylab = "log10(S(f))")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue", xlab = "log10(f)")
plot(f,log10(MTSE_fd3$spectrum[-1]), type = "l", main = "NW = 3", ylab = "log10(S(f))")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue", xlab = "log10(f)")
plot(f,log10(MTSE_fd4$spectrum[-1]), type = "l", main = "NW = 4", ylab = "log10(S(f))")
lines(log10(seq(0.001,0.5,length.out = 1025)),-2*d*log10(2*abs(sin(abs(pi*seq(0.005,0.5,length.out = 1025))))), col = "blue", xlab = "log10(f)")
```
## AVAR




## Power Law Processes

In clock frequency stability analysis typically physicists will assume a certain power law process and then calculate an Allan Deviation based on that assumption. With spectral analysis, however, we can test these assumptions, do the same kinds of analyses that clock physicists do, and more.


### Linear Drift
(Reappraisal 1987)

Due to environmental and systemic factors, there could be linear drifts in fractional frequency data $\{y_t\}$, so we actually have data of the form: 

$$y_t' = a + bt + y_t $$

According to Reappraisal, while the two sample variance has ways of dealing with this, namely, to estimate $b$ by:
 $$\hat{b} = \frac{y_N' - y_1'}{N-1} $$ where $N$ is the number of data points. Then estimate $\sigma_y^2(y,\tau)$ by $$ \hat{y_t} = y_t' - \hat{b}t$$. This yields a downwardly biased estimator of $\sigma_y^2$ (see Fig. 2 of Reappraisal).
 
If we have access to a direct spectral estimate of possibly tapered data, however, we can more effectively remove this linear drift using differencing of the data which, in the spectral realm, equates to an application of a simple linear filter. 


### Fitting a line for a Power Law process
(Multitaper for Power Law 98)

Clock Data is often assumed to be from a power law process, which has spectrum of the form $S(f) = h_\alpha f^\alpha$ for $|f| \leq \frac{1}{2}$.

The goal then is to estimate $\alpha$ using this baseline assumption about the spectral form of the underlying process.

#### Two Sample Variance Technique

There is a well-known result (according to Reappraisal), that the two-sample variance is well approximated by $\frac{A_\alpha}{\tau^{\alpha + 1}}$ for $A_\alpha$ dependent on $\alpha$, but not $h_\alpha$.

For reasons outlined in more detail in Reappraisal..., we may use the regression model:

 $$\log(\hat{\sigma}_y^2(2;2^k)) = \log(A_\alpha) - (\alpha + 1)k\log(2) + \eta_k\ $$
for $k = 0,\ldots,p-1$ where  where the $\eta_k$ are error terms that do not meet the assumptions of OLS due to heteroskedasticity and they are correlated--to name just two. This means we can estimate $\alpha$ 


#### Spectral Technique

Using a spectral estimate $\hat{S(f_k})$, we can find $\alpha$ by using the following regression:

$$\log(\hat{S}(f_k)) = \log(h_\alpha) + \alpha \log(f_k) $$ for $k = 1, ..., M$

where $M = N/2$ and $f_k = k/N$. Here we more closely match the assumptions of classical regression and thus we can obtain estimates for $\alpha$ and $h_\alpha$ as well as meaningful uncertainty estimates.

### White Noise Example

We recreate the analysis done in "Reappraisal..." where we generate 1000 WN(0,1) processes (a spectrum for which we know $\alpha = 0$) and try both techniques.

```{r, echo = FALSE}

#We already know for WN(0,1), S_y(f) = 1

#1. Create some data: 1000 realizations of N=128 WN(0,1)
N <- 128
Y <- matrix(NA, nrow = 1000,ncol = 128)

for(i in 1:1000){
set.seed(i)
Y[i,] <- rnorm(N,mean = 0, sd = 1)
}

#2. Calculate AVAR estimates
AVAR.Y <- matrix(NA, nrow = 1000, ncol = 6)
for(i in 1:1000){
  AVAR.Y[i,] <- getAvars(N,Y[i,],taus = 2^(0:5))$avarRes$avars
}

#3. Calculate S_y(f_k) estimates
Spec.Y <- matrix(NA, nrow = 1000, ncol = 64)
for(i in 1:1000){
  Spec.Y[i,] <- spec.pgram(Y[i,], plot = FALSE)$spec
}

#4. fit lines in log-log space to these estimates
v_k <- (0:5)*log(2)
alpha.ests_AVAR <- alpha.ests_spec <- rep(NA, times = 1000)

for(i in 1:1000){
 alpha.ests_AVAR[i] <-  -(lm(log(AVAR.Y[i,]) ~ v_k)$coefficients[2] + 1)
}

log_fk <- log(1:64/128)

for(i in 1:1000){
  alpha.ests_spec[i] <-  lm(log(Spec.Y[i,]) ~  log_fk)$coefficients[2] 
}


print(paste("AVAR alpha estimate: ", mean(alpha.ests_AVAR)))
print(paste("Spectrum alpha estimate: ", mean(alpha.ests_spec)))


```

This shows the bias of the two sample approach.

### Detection of Periodic Components
(Reappraisal 1987)



## Clock Data Example


```{r}
#load clock data
t <- clock_df$time
y <- clock_df$AlYb

plot(t,y)

hist(na.omit(y))
acf(na.omit(y)) 


```
This doesn't seem like white noise based on the acf, let's take a look at what the allan deviation might tell us.

```{r, echo = FALSE}

#calculate AVAR of X.t with missing gaps, pushed together
AVAR_missing <- getAvars(length(na.omit(y)),na.omit(y), taus = 2^(0:11)) #31903 data points

plot(log10(AVAR_missing$avarRes$taus), log10(AVAR_missing$avarRes$avars), xlab = "log10(tau)", ylab = "log10(AVAR)",pch = 19  , ylim = c(-4,1))

abline(a = 0, b = -1, col = "blue")

```


Spectral analysis:

```{r, echo = FALSE}

MTSE_AlYb <- multitaper_est(X.t = y[1:max(which(!is.na(y)))], NW = 3, K = 5)


```

```{r, echo = FALSE}


f <- log10(seq(0,0.5,length.out = length(MTSE_AlYb$spectrum)))[-1]

plot(f,log10(MTSE_AlYb$spectrum[-1]), type = "l", main = "Al/Yb clock data", ylab = "log10(S(f))")
abline(v = f[200])
```


Fit a line to the end piece of the spectrum?

```{r}
L <- length(MTSE_AlYb$spectrum)
lin.fit <- lm(log10(MTSE_AlYb$spectrum[200:L]) ~ f[200:L])

summary(lin.fit)
plot(lin.fit)
```

```{r}
plot(f,log10(MTSE_AlYb$spectrum[-1]), type = "l", main = "Al/Yb clock data", ylab = "log10(S(f))")
abline(a = lin.fit$coefficients[1], b = lin.fit$coefficients[2], col = "red")
```

## Bandpass Variance


A band-pass variance, defined: $$\beta_y^2(\tau) = 2\int_{1/4\tau}^{1/2\tau} S_y(f)df$$
according to "Reappraisal.." + Rutman 1978. For certain power-law processes $$\beta_y^2(\tau) \approx \frac{1}{2} \sigma^2_y(2;\tau)$$.

What would the estimate of bandpass variance be? Let $\tau = N/4$, where $N$ = length of the data set. Then numerically integrate our spectral estimate: $$2\beta_y^2(N) = 4\int_{1/4N}^{1/2N} \hat{S}_y(f)df$$


Below is an estimate using the clock data above:

```{r}
N = length(na.omit(clock_df$AlYb))
freqs <- seq(0,0.5, length.out = length(MTSE_AlYb$spectrum))
4*freqs[2]*MTSE_AlYb$spectrum[2]
```

### Bandpass Variance for White Noise

Just as a check, consider a white noise process. This has a flat spectrum, so the bandpass variance would be $$2a[\frac{1}{2\tau} - \frac{1}{4\tau}] = \frac{a}{2\tau} $$ where $a$ is the variance of the white noise process.




